---
title: "LMM"
output: 
  html_document:
    toc: true
    toc_depth: 5
    toc_float: True
date: '2024-08-02'
---

# load packages for Linear Mixed Model analysis
```{r setup, include=FALSE}
if(!require(rstatix)) install.packages("rstatix")
library(rstatix)

if(!require(readxl)) install.packages("readxl")
library(readxl)

if(!require(writexl)) install.packages("writexl")
library(writexl)

if(!require(lmerTest)) install.packages("lmerTest")
library(lmerTest)

if(!require(magrittr)) install.packages("magrittr")
library(magrittr)

if(!require(dplyr)) install.packages("dplyr")
library(dplyr)

if(!require(reshape2)) install.packages("reshape2")
library(reshape2)

```

# 3 Result (The following headings respond to the manuscript sections)
## 3.1 Manipulation checks and results
### 3.1.1 KSS ratings
```{r}
# read data
KSSdata <- read_excel("/Users/betty/Desktop/KSS(processed).xlsx", sheet=1)


# as factors
KSSdata$condition <- as.factor(KSSdata$condition)
KSSdata$`agegroup2` <- as.factor(KSSdata$`agegroup2`)
KSSdata$beforeafter <- as.factor(KSSdata$beforeafter)

# groupby
KSSdata %>%
  group_by(condition, `agegroup2`, beforeafter)


# younger drivers
filter.young<- filter(KSSdata, agegroup2 =="Y") 
## main effect of countermeasures for before countermeasure
filter1<- filter(filter.young, beforeafter =="before") 
res.fried.KSS.young.before <- filter1 %>% friedman_test(score ~ condition |sub)
res.fried.KSS.young.before
## post hoc analysis of countermeasures for before countermeasure
pwc.KSS.inter.young.before <- filter1 %>% wilcox_test(score ~ condition, paired = TRUE, p.adjust.method = "bonferroni")
pwc.KSS.inter.young.before
## main effect of countermeasures for after countermeasure
filter1<- filter(filter.young, beforeafter =="after") 
res.fried.KSS.young.after <- filter1 %>% friedman_test(score ~ condition |sub)
res.fried.KSS.young.after
## post hoc analysis of countermeasures for after countermeasure
pwc.KSS.inter.young.after <- filter1 %>% wilcox_test(score ~ condition, paired = TRUE, p.adjust.method = "bonferroni")
pwc.KSS.inter.young.after


# middle-aged drivers
filter.elder<- filter(KSSdata, agegroup2 =="M") 
## main effect of countermeasures for before countermeasure
filter1<- filter(filter.elder, beforeafter =="before") 
res.fried.KSS.middle.before <- filter1 %>% friedman_test(score ~ condition |sub)
res.fried.KSS.middle.before
## post hoc analysis of countermeasures for before countermeasure
pwc.KSS.inter.middle.before <- filter1 %>% wilcox_test(score ~ condition, paired = TRUE, p.adjust.method = "bonferroni")
pwc.KSS.inter.middle.before
## main effect of countermeasures for after countermeasure
filter1<- filter(filter.elder, beforeafter =="after") 
res.fried.KSS.middle.after <- filter1 %>% friedman_test(score ~ condition |sub)
res.fried.KSS.middle.after
## post hoc analysis of countermeasures for after countermeasure
pwc.KSS.inter.middle.after <- filter1 %>% wilcox_test(score ~ condition, paired = TRUE, p.adjust.method = "bonferroni")
pwc.KSS.inter.middle.after
```

### 3.1.2 PVT accuracy
```{r}
# read data
databutton <- read_excel("/Users/betty/Desktop/buttonprocessed.xlsx", sheet=1)
databutton  <- select(databutton,"sub","agegroup2","week","condition","order","beforeafter","score2")

# as factors
databutton$condition <- as.factor(databutton$condition)
databutton$beforeafter <- as.factor(databutton$beforeafter)
databutton$agegroup2 <- as.factor(databutton$agegroup2)

# Hypothesis testing
## QQ plot for residual
lmmodel <- lm(score2 ~ condition, databutton)
summary(lmmodel)
res_lmmodel <- residuals(lmmodel)
qqnorm(res_lmmodel) 
## KS test for residual
ks.test(res_lmmodel, "pnorm", mean(res_lmmodel), sd(res_lmmodel))
## Levene’s test for residual
library(car)
leveneTest(res_lmmodel, databutton$condition)


# younger drivers
filter.Y <- filter(databutton, agegroup2 =="Y") 
## main effect of countermeasures for before countermeasure
filter1.before <- filter(filter.Y, beforeafter =="beforeintervene") 
Model <- lmer(data = filter1.before, score2 ~condition +(1|sub) )
summary(Model)
anova(Model)
## post hoc analysis of countermeasures for before countermeasure
pwc.PVT.inter.young.before <- filter1.before %>% pairwise_t_test(score2 ~ condition, p.adjust.method = "bonferroni", paired = TRUE) 
pwc.PVT.inter.young.before 
## main effect of countermeasures for after countermeasure
filter1.after<- filter(filter.Y, beforeafter =="afterintervene") 
Model.PVT.2 <- lmer(data = filter1.after, score2~condition +(1|sub) )
summary(Model.PVT.2)
anova(Model.PVT.2)
## post hoc analysis of countermeasures for after countermeasure
pwc.PVT.inter.young.after <- filter1.after %>% pairwise_t_test(score2 ~ condition, p.adjust.method = "bonferroni", paired = TRUE) 
pwc.PVT.inter.young.after 

# middle-aged group
filter.M <- filter(databutton, agegroup2 =="M") 
## main effect of countermeasures for before countermeasure
filter1.before <- filter(filter.M, beforeafter =="beforeintervene") 
Model <- lmer(data = filter1.before, score2 ~condition +(1|sub) )
summary(Model)
anova(Model)
## post hoc analysis of countermeasures for before countermeasure
pwc.PVT.inter.middle.before <- filter1.before %>% pairwise_t_test(score2 ~ condition, p.adjust.method = "bonferroni", paired = TRUE) 
pwc.PVT.inter.middle.before
## main effect of countermeasures for after countermeasure
filter1.after<- filter(filter.M, beforeafter =="afterintervene") 
Model <- lmer(data = filter1.after, score2~condition +(1|sub) )
summary(Model)
anova(Model)
## post hoc analysis of countermeasures for after countermeasure
pwc.PVT.inter.middle.after <- filter1.after %>% pairwise_t_test(score2 ~ condition, p.adjust.method = "bonferroni", paired = TRUE) 
pwc.PVT.inter.middle.after
```

## 3.2 Driving performance 
### 3.2.1 SDLP
```{r, echo=FALSE}
# read data
SDLPdata <- read_excel("/Users/betty/Desktop/Objectivedata(processed).xlsx", sheet=1)
SDLPdata  <- dplyr::select(SDLPdata ,"sub","age","agegroup2","condition","order","beforeafter", "SDLP")

# as factors
SDLPdata$condition <- as.factor(SDLPdata$condition)
SDLPdata$beforeafter <- as.factor(SDLPdata$beforeafter)
SDLPdata$agegroup2 <- as.factor(SDLPdata$agegroup2)

# Hypothesis testing
## QQ plot for residual
lmmodel <- lm(SDLP ~ condition*beforeafter, SDLPdata)
summary(lmmodel)
res_lmmodel <- residuals(lmmodel)
qqnorm(res_lmmodel)
## KS test for residual
ks.test(res_lmmodel, "pnorm", mean(res_lmmodel), sd(res_lmmodel))
## Levene's test for residual
library(car)
leveneTest(lmmodel)

## QQ plot for residual (logged data)
lmmodel <- lm(log(SDLP) ~ condition*beforeafter, SDLPdata)
summary(lmmodel)
res_lmmodel <- residuals(lmmodel)
qqnorm(res_lmmodel)
## KS test for residual (logged data)
ks.test(res_lmmodel, "pnorm", mean(res_lmmodel), sd(res_lmmodel))
## Levene's test for residual (logged data)
library(car)
leveneTest(lmmodel)

# used logged data
SDLPdata$SDLP <- log(SDLPdata$SDLP)

# Linear Mixed Model output
Model <- lmer(data = SDLPdata, SDLP~condition*beforeafter*agegroup2+(1|sub))
summary(Model)
SDLP <- anova(Model)
SDLP

## the simple effect analysis for the interaction effect of countermeasure type and countermeasure timing
### comparisons of countermeasure types before countermeasure
filter1.before<- filter(SDLPdata, beforeafter  =="beforeintervene") 
Model <- lmer(data = filter1.before, SDLP~condition+(1|sub))
summary(Model)
anova(Model)
pwc.SDLP.inter.before <- filter1.before %>% pairwise_t_test(SDLP ~ condition, p.adjust.method = "bonferroni", paired = TRUE) 
pwc.SDLP.inter.before
### comparisons of countermeasure types during countermeasure
filter1.on<- filter(SDLPdata, beforeafter  =="on") 
Model <- lmer(data = filter1.on, SDLP~condition+(1|sub))
summary(Model)
anova(Model)
pwc.SDLP.inter.during <- filter1.on %>% pairwise_t_test(SDLP ~ condition, p.adjust.method = "bonferroni", paired = TRUE) 
pwc.SDLP.inter.during
```

### 3.2.2 SDVHA
```{r, echo=FALSE}
# read data
SDVHdata <- read_excel("/Users/betty/Desktop/Objectivedata(processed).xlsx", sheet=1)
SDVHdata  <- dplyr::select(SDVHdata ,"sub","age","agegroup2","condition","order","beforeafter", "SDVH")

# as factors
SDVHdata$condition <- as.factor(SDVHdata$condition)
SDVHdata$beforeafter <- as.factor(SDVHdata$beforeafter)
SDVHdata$agegroup2 <- as.factor(SDVHdata$agegroup2)

# Hypothesis testing
## QQ plot for residual
lmmodel <- lm(SDVH ~ condition*beforeafter, SDVHdata)
summary(lmmodel)
res_lmmodel <- residuals(lmmodel)
qqnorm(res_lmmodel)
## KS test for residual
ks.test(res_lmmodel, "pnorm", mean(res_lmmodel), sd(res_lmmodel))
## Levene's test for residual
leveneTest(lmmodel)

## QQ plot for residual (logged data)
lmmodel <- lm(log(SDVH) ~ condition*beforeafter, SDVHdata)
summary(lmmodel)
res_lmmodel <- residuals(lmmodel)
qqnorm(res_lmmodel)
## KS test for residual (logged data)
ks.test(res_lmmodel, "pnorm", mean(res_lmmodel), sd(res_lmmodel))
## Levene's test for residual (logged data)
leveneTest(lmmodel)

# used logged data
SDVHdata$SDVH <- log(SDVHdata$SDVH)

# Linear Mixed Model output
Model <- lmer(data = SDVHdata, SDVH~condition*beforeafter*agegroup2+(1|sub))
summary(Model)
SDVH <- anova(Model)
SDVH

## the simple effect analysis for the interaction effect of countermeasure type and countermeasure timing
### comparisons of countermeasure types before countermeasure
filter1.before<- filter(SDVHdata, beforeafter  =="beforeintervene") 
Model <- lmer(data = filter1.before, SDVH~condition+(1|sub))
summary(Model)
anova(Model)
pwc.SDVH.inter.before <- filter1.before %>% pairwise_t_test(SDVH ~ condition, p.adjust.method = "bonferroni", paired = TRUE) 
pwc.SDVH.inter.before
### comparisons of countermeasure types during countermeasure
filter1.on<- filter(SDVHdata, beforeafter  =="on") 
Model <- lmer(data = filter1.on, SDVH~condition+(1|sub))
summary(Model)
anova(Model)
pwc.SDVH.inter.during <- filter1.on %>% pairwise_t_test(SDVH ~ condition, p.adjust.method = "bonferroni", paired = TRUE) 
pwc.SDVH.inter.during
```

### 3.2.3 SDspeed
```{r, echo=FALSE}
# read data
SDSpeeddata <- read_excel("/Users/betty/Desktop/Objectivedata(processed).xlsx", sheet=1)
SDSpeeddata  <- dplyr::select(SDSpeeddata,"sub","age","agegroup2","condition","order","beforeafter", "SDSpeed")

# as factors
SDSpeeddata$condition <- as.factor(SDSpeeddata$condition)
SDSpeeddata$beforeafter <- as.factor(SDSpeeddata$beforeafter)
SDSpeeddata$agegroup2 <- as.factor(SDSpeeddata$agegroup2)

# Hypothesis testing
## QQ plot for residual
lmmodel <- lm(SDSpeed ~ condition*beforeafter, SDSpeeddata)
summary(lmmodel)
res_lmmodel <- residuals(lmmodel)
qqnorm(res_lmmodel)
## KS test for residual
ks.test(res_lmmodel, "pnorm", mean(res_lmmodel), sd(res_lmmodel))
## Levene's test for residual
leveneTest(lmmodel)

## QQ plot for residual (logged data)
lmmodel <- lm(log(SDSpeed) ~ condition*beforeafter, SDSpeeddata)
summary(lmmodel)
res_lmmodel <- residuals(lmmodel)
qqnorm(res_lmmodel)
## KS test for residual (logged data)
ks.test(res_lmmodel, "pnorm", mean(res_lmmodel), sd(res_lmmodel))
## Levene's test for residual (logged data)
leveneTest(lmmodel)

# use logged data
SDSpeeddata$SDSpeed <- log(SDSpeeddata$SDSpeed)

# Linear Mixed Model output
Model <- lmer(data = SDSpeeddata, SDSpeed~condition*beforeafter*agegroup2+(1|sub))
summary(Model)
SDspeed <- anova(Model)
SDspeed

## the post hoc analysis for the main effect of countermeasure timing
pwc.SDSpeed.timing <- SDSpeeddata %>% pairwise_t_test(SDSpeed ~ beforeafter, p.adjust.method = "bonferroni", paired = TRUE) 
pwc.SDSpeed.timing

## the simple effect analysis for the interaction effect of countermeasure type and countermeasure timing
### comparisons of countermeasure types before countermeasure
filter.before<- filter(SDSpeeddata, beforeafter  =="beforeintervene") 
Model <- lmer(data = filter.before, SDSpeed~condition+(1|sub))
summary(Model)
anova(Model)
pwc.SDSpeed.inter.before <- filter.before %>% pairwise_t_test(SDSpeed~ condition, p.adjust.method = "bonferroni", paired = TRUE) 
pwc.SDSpeed.inter.before
### comparisons of countermeasure types during countermeasure
filter.during<- filter(SDSpeeddata, beforeafter  =="on") 
Model <- lmer(data = filter.during, SDSpeed~condition+(1|sub))
summary(Model)
anova(Model)
pwc.SDSpeed.inter.during <- filter.during %>% pairwise_t_test(SDSpeed ~ condition, p.adjust.method = "bonferroni", paired = TRUE) 
pwc.SDSpeed.inter.during
```

## 3.3 Physiological indicators
### 3.3.1 SCL (skin conductance level extracted from electrodermal activity (EDA))
```{r, echo=FALSE}
# read data
SCLdata <- read_excel("/Users/betty/Desktop/Objectivedata(processed).xlsx", sheet=1)
SCLdata  <- dplyr::select(SCLdata ,"sub","age","agegroup2","condition","order","beforeafter", "SCL")

# as factors
SCLdata$condition <- as.factor(SCLdata$condition)
SCLdata$beforeafter <- as.factor(SCLdata$beforeafter)
SCLdata$agegroup2 <- as.factor(SCLdata$agegroup2)

# Hypothesis testing
## QQ plot for residual
lmmodel <- lm(SCL ~ condition*beforeafter, SCLdata)
summary(lmmodel)
res_lmmodel <- residuals(lmmodel)
qqnorm(res_lmmodel)
## KS test for residual
ks.test(res_lmmodel, "pnorm", mean(res_lmmodel), sd(res_lmmodel))
## Levene's test for residual
library(car)
leveneTest(lmmodel)

# Linear Mixed Model output
Model <- lmer(data = SCLdata, SCL~condition*beforeafter*agegroup2+(1|sub))
summary(Model)
SCL <- anova(Model)
SCL

## the post hoc analysis for the main effect of countermeasure type
pwc.SCL.condition <- SCLdata %>% pairwise_t_test(SCL ~ condition, p.adjust.method = "bonferroni", paired = TRUE) 
pwc.SCL.condition

## the post hoc analysis for the main effect of countermeasure timing
pwc.SCL.timing <- SCLdata %>% pairwise_t_test(SCL ~ beforeafter, p.adjust.method = "bonferroni", paired = TRUE) 
pwc.SCL.timing

## the simple effect analysis for the interaction effect of countermeasure type and countermeasure timing
### comparisons of countermeasure types before countermeasure
filter1.before<- filter(SCLdata, beforeafter  =="beforeintervene") 
Model <- lmer(data = filter1.before, SCL~condition+(1|sub))
summary(Model)
anova(Model)
pwc.SCL.inter.before <- filter1.before %>% pairwise_t_test(SCL ~ condition, p.adjust.method = "bonferroni", paired = TRUE) 
pwc.SCL.inter.before
### comparisons of countermeasure types during countermeasure
filter1.on<- filter(SCLdata, beforeafter  =="on") 
anova_test(filter1.on, wid = sub, within = c(condition), dv = SCL)
Model <- lmer(data = filter1.on, SCL~condition+(1|sub))
summary(Model)
anova(Model)
pwc.SCL.inter.during<- filter1.on %>% pairwise_t_test(SCL ~ condition, p.adjust.method = "bonferroni", paired = TRUE) 
pwc.SCL.inter.during
```


### 3.3.2 SDNN (Standard Deviation of Normal-to-Normal intervals (SDNN) extracted from photoplethysmography (PPG))
```{r, echo=FALSE}
# read data
SDNNdata <- read_excel("/Users/betty/Desktop/Objectivedata(processed).xlsx", sheet=1)
SDNNdata  <- dplyr::select(SDNNdata ,"sub","age","agegroup2","condition","order","beforeafter", "SDNN")

# as factors
SDNNdata$condition <- as.factor(SDNNdata$condition)
SDNNdata$beforeafter <- as.factor(SDNNdata$beforeafter)
SDNNdata$agegroup2 <- as.factor(SDNNdata$agegroup2)

# Hypothesis testing
## QQ plot for residual
lmmodel <- lm(SDNN ~ condition*beforeafter, SDNNdata)
summary(lmmodel)
res_lmmodel <- residuals(lmmodel)
qqnorm(res_lmmodel)
## KS test for residual
ks.test(res_lmmodel, "pnorm", mean(res_lmmodel), sd(res_lmmodel))
## Levene's test for residual
leveneTest(lmmodel)

## QQ plot for residual (logged data)
lmmodel <- lm(log(SDNN) ~ condition*beforeafter, SDNNdata)
summary(lmmodel)
res_lmmodel <- residuals(lmmodel)
qqnorm(res_lmmodel)
## KS test for residual (logged data)
ks.test(res_lmmodel, "pnorm", mean(res_lmmodel), sd(res_lmmodel))
## Levene's test for residual (logged data)
leveneTest(lmmodel)


# use logged data
SDNNdata$SDNN <- log (SDNNdata$SDNN)

# Linear Mixed Model output
Model <- lmer(data = SDNNdata, SDNN~condition*beforeafter*agegroup2+(1|sub))
summary(Model)
anova(Model)
SDNN <- anova(Model)
SDNN

## the post hoc analysis for the main effect of countermeasure timing
pwc.SDNN.timing <- SDNNdata %>% pairwise_t_test(SDNN ~ beforeafter, p.adjust.method = "bonferroni", paired = TRUE) 
pwc.SDNN.timing


## the simple effect analysis for the interaction effect of countermeasure type and countermeasure timing
### comparisons of countermeasure types before countermeasure
filter1.before<- filter(SDNNdata, beforeafter  =="beforeintervene") 
Model <- lmer(data = filter1.before, SDNN~condition+(1|sub))
summary(Model)
anova(Model)
pwc.SDNN.inter.before <- filter1.before %>% pairwise_t_test(SDNN ~ condition, p.adjust.method = "bonferroni", paired = TRUE) 
pwc.SDNN.inter.before
### comparisons of countermeasure types during countermeasure
filter1.on<- filter(SDNNdata, beforeafter  =="on") 
Model <- lmer(data = filter1.on, SDNN~condition+(1|sub))
summary(Model)
anova(Model)
pwc.SDNN.inter.during <- filter1.on %>% pairwise_t_test(SDNN ~ condition, p.adjust.method = "bonferroni", paired = TRUE) 
pwc.SDNN.inter.during
```

### 3.3.3 Breathing rate
```{r, echo=FALSE}
## read data
RESPdata <- read_excel("/Users/betty/Desktop/Objectivedata(processed).xlsx", sheet=1)
RESPdata  <- dplyr::select(RESPdata ,"sub","age","agegroup2","condition","order","beforeafter", "RESP")

# as factors
RESPdata$condition <- as.factor(RESPdata$condition)
RESPdata$beforeafter <- as.factor(RESPdata$beforeafter)
RESPdata$agegroup2 <- as.factor(RESPdata$agegroup2)

# Hypothesis testing
## QQ plot for residual
lmmodel <- lm(RESP ~ condition*beforeafter, RESPdata)
summary(lmmodel)
res_lmmodel <- residuals(lmmodel)
qqnorm(res_lmmodel)
## KS test for residual
ks.test(res_lmmodel, "pnorm", mean(res_lmmodel), sd(res_lmmodel))
## Levene's test for residual
leveneTest(lmmodel)

## QQ plot for residual (logged data)
lmmodel <- lm(log(RESP) ~ condition*beforeafter, RESPdata)
summary(lmmodel)
res_lmmodel <- residuals(lmmodel)
qqnorm(res_lmmodel)
## KS test for residual (logged data)
ks.test(res_lmmodel, "pnorm", mean(res_lmmodel), sd(res_lmmodel))
## Levene's test for residual (logged data)
leveneTest(lmmodel)

# use logged data
RESPdata$RESP <- log(RESPdata$RESP)

# Linear Mixed Model output
Model <- lmer(data = RESPdata, RESP~condition*beforeafter*agegroup2+(1|sub))
summary(Model)
RESP <- anova(Model)
RESP
```


## 3.4 Eye movement indicators
### 3.4.1 Pupil diameter (PD)
```{r, echo=FALSE}
# read data
PDdata <- read_excel("/Users/betty/Desktop/Objectivedata(processed).xlsx", sheet=1)
PDdata  <- dplyr::select(PDdata ,"sub","age","agegroup2","condition","order","beforeafter", "PD")

# as factors
PDdata$condition <- as.factor(PDdata$condition)
PDdata$beforeafter <- as.factor(PDdata$beforeafter)
PDdata$agegroup2 <- as.factor(PDdata$agegroup2)

# Hypothesis testing
## QQ plot for residual
lmmodel <- lm(PD ~ condition*beforeafter, PDdata)
summary(lmmodel)
res_lmmodel <- residuals(lmmodel)
qqnorm(res_lmmodel)
## KS test for residual
ks.test(res_lmmodel, "pnorm", mean(res_lmmodel), sd(res_lmmodel))
## Levene's test for residual
leveneTest(lmmodel)

# Linear Mixed Model output
Model <- lmer(data = PDdata, PD~agegroup2*condition*beforeafter+(1|sub))
summary(Model)
PD <- anova(Model)
PD
## the post hoc analysis for the main effect of countermeasure type
pwc.PD.condition <- PDdata %>% pairwise_t_test(PD ~ condition, p.adjust.method = "bonferroni", paired = TRUE) 
pwc.PD.condition

## the post hoc analysis for the main effect of countermeasure timing
pwc.PD.timing <- PDdata %>% pairwise_t_test(PD ~ beforeafter, p.adjust.method = "bonferroni", paired = TRUE) 
pwc.PD.timing

## the simple effect analysis for the interaction effect of countermeasure type and countermeasure timing
### comparisons of countermeasure types before countermeasure
filter1.before<- filter(PDdata, beforeafter  =="beforeintervene") 
Model <- lmer(data = filter1.before, PD~condition+(1|sub))
summary(Model)
anova(Model)
pwc.PD.inter.before <- filter1.before %>% pairwise_t_test(PD ~ condition, p.adjust.method = "bonferroni", paired = TRUE) 
pwc.PD.inter.before

### comparisons of countermeasure types during countermeasure
filter1.on<- filter(PDdata, beforeafter  =="on") 
Model <- lmer(data = filter1.on, PD~condition+(1|sub))
summary(Model)
anova(Model)
pwc.PD.inter.during <- filter1.on %>% pairwise_t_test(PD ~ condition, p.adjust.method = "bonferroni", paired = TRUE) 
pwc.PD.inter.during 
```

### 3.4.2 PVRC (proportion of visit duration on the road centre)
```{r, echo=FALSE}
## read data
PVRCdata <- read_excel("/Users/betty/Desktop/Objectivedata(processed).xlsx", sheet=1)
PVRCdata  <- dplyr::select(PVRCdata ,"sub","age","agegroup2","condition","order","beforeafter", "PVRC")

# as factors
PVRCdata$condition <- as.factor(PVRCdata$condition)
PVRCdata$beforeafter <- as.factor(PVRCdata$beforeafter)
PVRCdata$agegroup2 <- as.factor(PVRCdata$agegroup2)

# Hypothesis testing
## QQ plot for residual
lmmodel <- lm(PVRC ~ condition*beforeafter, PVRCdata)
summary(lmmodel)
res_lmmodel <- residuals(lmmodel)
qqnorm(res_lmmodel)
## KS test for residual
ks.test(res_lmmodel, "pnorm", mean(res_lmmodel), sd(res_lmmodel))
## Levene's test for residual
leveneTest(lmmodel)

# Linear Mixed Model output
Model <- lmer(data = PVRCdata, PVRC~condition*beforeafter*agegroup2+(1|sub))
summary(Model)
PVRC <- anova(Model)
PVRC
## the post hoc analysis for the main effect of countermeasure timing
pwc.PVRC.timing <- PVRCdata %>% pairwise_t_test(PVRC ~ beforeafter, p.adjust.method = "bonferroni", paired = TRUE) 
pwc.PVRC.timing

## the simple effect analysis for the interaction effect of countermeasure type and countermeasure timing
### comparisons of countermeasure types before countermeasure
filter1.before<- filter(PVRCdata, beforeafter  =="beforeintervene") 
Model <- lmer(data = filter1.before, PVRC~condition+(1|sub))
summary(Model)
anova(Model)
pwc.PVRC.inter.before <- filter1.before %>% pairwise_t_test(PVRC ~ condition, p.adjust.method = "bonferroni", paired = TRUE) 
pwc.PVRC.inter.before
### comparisons of countermeasure types during countermeasure
filter1.on<- filter(PVRCdata, beforeafter  =="on") 
Model <- lmer(data = filter1.on, PVRC~condition+(1|sub))
summary(Model)
anova(Model)
pwc.PVRC.inter.during <- filter1.on %>% pairwise_t_test(PVRC ~ condition, p.adjust.method = "bonferroni", paired = TRUE) 
pwc.PVRC.inter.during
```

### 3.4.3 Blink rate (BL)
```{r, echo=FALSE}
# read data
BLdata <- read_excel("/Users/betty/Desktop/Objectivedata(processed).xlsx", sheet=1)
BLdata  <- dplyr::select(BLdata ,"sub","age","agegroup2","condition","order","beforeafter", "BL")

# as factors
BLdata$condition <- as.factor(BLdata$condition)
BLdata$beforeafter <- as.factor(BLdata$beforeafter)
BLdata$agegroup2 <- as.factor(BLdata$agegroup2)

# Hypothesis testing
## QQ plot for residual
lmmodel <- lm(BL ~ condition*beforeafter, BLdata)
summary(lmmodel)
res_lmmodel <- residuals(lmmodel)
qqnorm(res_lmmodel)
## KS test for residual
ks.test(res_lmmodel, "pnorm", mean(res_lmmodel), sd(res_lmmodel))
## Levene's test for residual
leveneTest(lmmodel)

## QQ plot for residual (logged data)
lmmodel <- lm(log(BL) ~ condition*beforeafter, BLdata)
summary(lmmodel)
res_lmmodel <- residuals(lmmodel)
qqnorm(res_lmmodel)
## KS test for residual (logged data)
ks.test(res_lmmodel, "pnorm", mean(res_lmmodel), sd(res_lmmodel))
## Levene's test for residual (logged data)

# use logged data
BLdata$BL <- log(BLdata$BL)

# Linear Mixed Model output
Model <- lmer(data = BLdata, BL~condition*beforeafter*agegroup2+(1|sub))
summary(Model)
anova(Model)
BL <- anova(Model)
BL
## the post hoc analysis for the main effect of countermeasure timing
pwc.BL.condition <- BLdata %>% pairwise_t_test(BL ~ condition, p.adjust.method = "bonferroni", paired = TRUE) 
pwc.BL.condition

## the simple effect analysis for the interaction effect of countermeasure type and countermeasure timing
### comparisons of countermeasure types before countermeasure
filter1.before<- filter(BLdata, beforeafter  =="beforeintervene") 
Model <- lmer(data = filter1.before, BL~condition+(1|sub))
summary(Model)
anova(Model)
pwc.BL.inter.before <- filter1.before %>% pairwise_t_test(BL ~ condition, p.adjust.method = "bonferroni", paired = TRUE) 
pwc.BL.inter.before
### comparisons of countermeasure types during countermeasure
filter1.on<- filter(BLdata, beforeafter  =="on") 
Model <- lmer(data = filter1.on, BL~condition+(1|sub))
summary(Model)
anova(Model)
pwc.BL.inter.during <- filter1.on %>% pairwise_t_test(BL ~ condition, p.adjust.method = "bonferroni", paired = TRUE) 
pwc.BL.inter.during
```

## 3.5 Heterogeneity 
### Packages for calculating Heterogeneity 
```{r}
if(!require(here)) install.packages("here")
library(here)

if(!require(ggplot2)) install.packages("ggplot2")
library(ggplot2)

if(!require(dplyr)) install.packages("dplyr")
library(dplyr)

if(!require(tidyr)) install.packages("tidyr")
library(tidyr)

if(!require(viridis)) install.packages("viridis")
library(viridis)

if(!require(gridExtra)) install.packages("gridExtra")
library(gridExtra)

if(!require(readr)) install.packages("readr")
library(readr)

if(!require(plyr)) install.packages("plyr")
library(plyr)

if(!require(stringr)) install.packages("stringr")
library(stringr)

if(!require(readxl)) install.packages("readxl")
library(readxl)

if(!require(data.table)) install.packages("data.table")
library(data.table)

if(!require(lmerTest)) install.packages("lmerTest")
library(lmerTest)

if(!require(plotly)) install.packages("plotly")
library(plotly)

if(!require(zoo)) install.packages("zoo")
library(zoo)

if(!require(viridis)) install.packages("viridis")
library(viridis)

if(!require(brms)) install.packages("brms")
library(brms)

if(!require(purrr)) install.packages("purrr")
library(purrr)

if(!require(MASS)) install.packages("MASS")
library(MASS)

if(!require(rstan)) install.packages("rstan")
library(rstan)

if(!require(ggdist)) install.packages("ggdist")
library(ggdist)

if(!require(bayestestR)) install.packages("bayestestR")
library(bayestestR)

if(!require(posterior)) install.packages("posterior")
library(posterior)

if(!require(distributional)) install.packages("distributional")
library(distributional)

if(!require(cowplot)) install.packages("cowplot")
library(cowplot)

if(!require(modelr)) install.packages("modelr")
library(modelr)

if(!require(purrr)) install.packages("purrr")
library(purrr)

if(!require(forcats)) install.packages("forcats")
library(forcats)

if(!require(tidybayes)) install.packages("tidybayes")
library(tidybayes)

if(!require(bayesplot)) install.packages("bayesplot")
library(bayesplot)

if(!require(BayesFactor)) install.packages("BayesFactor")
library(BayesFactor)

if(!require(patchwork)) install.packages("patchwork")
library(patchwork)

if(!require(scales)) install.packages("scales")
library(scales)
```

### SDLP
```{r}
# scale the data
SDLPdata$age.c <- scale(SDLPdata$age, center = T, scale = F)

# fit model
Model <- lmer(data = SDLPdata, SDLP~beforeafter*condition*age.c+(1+beforeafter|sub))
summary(Model)

# calculate the contribution of age 
# Get interaction coefficient between age and beforeafter
bXa.coeff <- fixef(Model)[12] 

# Random effect in model for beforeafter - squared to compute variance 
residvalvar <- 0.00361

# Calculate the implied total random effect variance - interaction squared multiplied by variance in age + residual variance.
imptotalvalvar <- bXa.coeff^2*sd(SDLPdata$age.c)^2 + residvalvar

# Variance in mu difference explained by age
V_SDLP <-1- (residvalvar/imptotalvalvar)
V_SDLP

# plot the data
## computing random effects 
ageranef_original <- as.data.frame(ranef(Model)) %>% 
  dplyr::select("grp","term","condval")

library(reshape2)
ageranef <- dcast(ageranef_original, 
                  grp~term, 
                  timevar = c("condval")) 

ageranef <- ageranef %>% rename("intercept_age"="(Intercept)","slope_age"="beforeafteron","sub"="grp")


## Create dataset with one line per person with mileage score 
ageranef <- merge(ageranef, SDLPdata, by = 'sub')
ageranef <- ageranef %>%
  dplyr::group_by(sub) %>%
  dplyr::slice(1) %>%
  dplyr::select(sub, age.c, intercept_age, slope_age)

## Person-specific implied total n-back effects for model accounting for age
ranef.age.pred.tot <- fixef(Model)[2] + # fixed effect for n-back
  fixef(Model)[12] * ageranef$age.c +   # fixed effect for n-back X age interaction # scale of age 
  ageranef$slope_age # estimate random effect of slopes

## relationship between the effect of N-back and age
age.nback.effect.sge_SDLP <- ggplot(ageranef, aes(age.c, ranef.age.pred.tot)) + 
  geom_vline(xintercept = 0, size = .5, color = "black", linetype="solid") +
  geom_hline(yintercept = mean(ranef.age.pred.tot), size = .5, color = "black", linetype="solid") +
  geom_jitter(height = 0, width = 0, size = 4,
              shape = 21, colour = "black", fill = viridis(5)[3], alpha = .95, stroke = 1) + 
  stat_smooth(method = 'lm', ci = T, color = viridis(5)[3], size = 1) + 
  xlab("Age (mean centered)") +
  ylab(expression("Implied total heterogeneity of "~ SDLP)) +
  #ylim(-.15, .1) +
  theme_bw() +
  theme(legend.position = "bottom", legend.direction = "horizontal", axis.title.x = element_text(size = 11), axis.text.x = element_text(size = 15), axis.title.y = element_text(size = 11), axis.text.y = element_text(size = 15), title = element_text(size = 18), legend.title = element_text(size = 12), legend.text = element_text(size = 12), legend.key = element_blank(), legend.key.width = unit(0.3, 'cm'), legend.key.size = unit(0.1, 'cm'), plot.title = element_text(size = 7.5, face = "bold"), strip.text = element_text(face = "bold", size = 10))
age.nback.effect.sge_SDLP
```

### SDVHA
```{r}
# scale the data
SDVHdata$age.c <- scale(SDVHdata$age, center = T, scale = F)

# fit model
Model <- lmer(data = SDVHdata, SDVH~beforeafter*condition*age.c+(1+beforeafter|sub))
summary(Model)

# calculate the contribution of age 
# Get interaction coefficient between age and beforeafter
bXa.coeff <- fixef(Model)[12] 

# Random effect in model for beforeafter - squared to compute variance 
residvalvar <- 0.0003094

# Calculate the implied total random effect variance - interaction squared multiplied by variance in age + residual variance.
imptotalvalvar <- bXa.coeff^2*sd(SDVHdata$age.c)^2 + residvalvar

# Variance in mu difference explained by age
V_SDVH <- 1 - (residvalvar/imptotalvalvar)
V_SDVH

# plot the data
## computing random effects 
ageranef_original <- as.data.frame(ranef(Model)) %>% 
  dplyr::select("grp","term","condval")

library(reshape2)
ageranef <- dcast(ageranef_original, 
                  grp~term, 
                  timevar = c("condval")) 

ageranef <- ageranef %>% rename("intercept_age"="(Intercept)","slope_age"="beforeafteron","sub"="grp")


#Create dataset with one line per person with mileage score 
ageranef <- merge(ageranef, SDVHdata, by = 'sub')
ageranef <- ageranef %>%
  dplyr::group_by(sub) %>%
  dplyr::slice(1) %>%
  dplyr::select(sub, age.c, intercept_age, slope_age)



# Person-specific implied total n-back effects for model accounting for age
ranef.age.pred.tot <- fixef(Model)[2] + # fixed effect for n-back
  fixef(Model)[12] * ageranef$age.c +   # fixed effect for n-back X age interaction # scale of age 
  ageranef$slope_age # estimate random effect of slopes



# relationship between the effect of N-back and age
age.nback.effect.sge_SDVH <- ggplot(ageranef, aes(age.c, ranef.age.pred.tot)) + 
  geom_vline(xintercept = 0, size = .5, color = "black", linetype="solid") +
  geom_hline(yintercept = mean(ranef.age.pred.tot), size = .5, color = "black", linetype="solid") +
  geom_jitter(height = 0, width = 0, size = 4,
              shape = 21, colour = "black", fill = viridis(5)[3], alpha = .95, stroke = 1) + 
  stat_smooth(method = 'lm', ci = T, color = viridis(5)[3], size = 1) + 
  xlab("Age (mean centered)") +
  ylab(expression("Implied total heterogeneity of "~ SDVH)) +
  #ylim(-.15, .1) +
  theme_bw() +
  theme(legend.position = "bottom", legend.direction = "horizontal", axis.title.x = element_text(size = 11), axis.text.x = element_text(size = 15), axis.title.y = element_text(size = 11), axis.text.y = element_text(size = 15), title = element_text(size = 18), legend.title = element_text(size = 12), legend.text = element_text(size = 12), legend.key = element_blank(), legend.key.width = unit(0.3, 'cm'), legend.key.size = unit(0.1, 'cm'), plot.title = element_text(size = 7.5, face = "bold"), strip.text = element_text(face = "bold", size = 10))
age.nback.effect.sge_SDVH
```

### SDspeed
```{r}
# scale the data
SDSpeeddata$age.c <- scale(SDSpeeddata$age, center = T, scale = F)

# fit model
Model <- lmer(data = SDSpeeddata, SDSpeed~beforeafter*condition*age.c+(1+beforeafter|sub))
summary(Model)

# calculate the contribution of age 
# Get interaction coefficient between age and beforeafter
bXa.coeff <- fixef(Model)[12] 

# Random effect in model for beforeafter - squared to compute variance 
residvalvar <- 0.02416

# Calculate the implied total random effect variance - interaction squared multiplied by variance in age + residual variance.
imptotalvalvar <- bXa.coeff^2*sd(SDSpeeddata$age.c)^2 + residvalvar


# Variance in mu difference explained by age
V_SD_speed<- 1 - (residvalvar/imptotalvalvar)
V_SD_speed

# plot the data
## computing random effects 
ageranef_original <- as.data.frame(ranef(Model)) %>% 
  dplyr::select("grp","term","condval")

library(reshape2)
ageranef <- dcast(ageranef_original, 
                  grp~term, 
                  timevar = c("condval")) 

ageranef <- ageranef %>% rename("intercept_age"="(Intercept)","slope_age"="beforeafteron","sub"="grp")


## Create dataset with one line per person with mileage score 
ageranef <- merge(ageranef, SDSpeeddata, by = 'sub')
ageranef <- ageranef %>%
  dplyr::group_by(sub) %>%
  dplyr::slice(1) %>%
  dplyr::select(sub, age.c, intercept_age, slope_age)

## Person-specific implied total n-back effects for model accounting for age
ranef.age.pred.tot <- fixef(Model)[2] + # fixed effect for n-back
  fixef(Model)[12] * ageranef$age.c +   # fixed effect for n-back X age interaction # scale of age 
  ageranef$slope_age # estimate random effect of slopes

## relationship between the effect of N-back and age
age.nback.effect.sge_SDSpeed <- ggplot(ageranef, aes(age.c, ranef.age.pred.tot)) + 
  geom_vline(xintercept = 0, size = .5, color = "black", linetype="solid") +
  geom_hline(yintercept = mean(ranef.age.pred.tot), size = .5, color = "black", linetype="solid") +
  geom_jitter(height = 0, width = 0, size = 4,
              shape = 21, colour = "black", fill = viridis(5)[3], alpha = .95, stroke = 1) + 
  stat_smooth(method = 'lm', ci = T, color = viridis(5)[3], size = 1) + 
  xlab("Age (mean centered)") +
  ylab(expression("Implied total heterogeneity of "~ SD-speed)) +
  #ylim(-.15, .1) +
  theme_bw() +
  theme(legend.position = "bottom", legend.direction = "horizontal", axis.title.x = element_text(size = 11), axis.text.x = element_text(size = 15), axis.title.y = element_text(size = 11), axis.text.y = element_text(size = 15), title = element_text(size = 18), legend.title = element_text(size = 12), legend.text = element_text(size = 12), legend.key = element_blank(), legend.key.width = unit(0.3, 'cm'), legend.key.size = unit(0.1, 'cm'), plot.title = element_text(size = 7.5, face = "bold"), strip.text = element_text(face = "bold", size = 10))
age.nback.effect.sge_SDSpeed
```

### SCL
```{r}
# scale the data
SCLdata$age.c <- scale(SCLdata$age, center = T, scale = F)

# fit model
Model <- lmer(data = SCLdata, SCL~beforeafter*condition+(1+beforeafter|sub))
summary(Model)

# calculate the contribution of age 
# Get interaction coefficient between age and beforeafter
bXa.coeff <- fixef(Model)[12] # the estimate parameter of age.c and beforeafter interaction effect

#Random effect in model for beforeafter - squared to compute variance 
residvalvar <- 0.05661

#Calculate the implied total random effect variance - interaction squared multiplied by variance in age + residual variance.
imptotalvalvar <- bXa.coeff^2*sd(SCLdata$age.c)^2 + residvalvar

## Variance in mu difference explained by age 
V_SCL <- 1 - (residvalvar/imptotalvalvar)
V_SCL

# plot the data
# computing random effects 
ageranef_original <- as.data.frame(ranef(Model)) %>% 
  dplyr::select("grp","term","condval")

library(reshape2)
ageranef <- dcast(ageranef_original, 
                  grp~term, 
                  timevar = c("condval")) 

ageranef <- ageranef %>% rename("intercept_age"="(Intercept)","slope_age"="beforeafteron","sub"="grp")


#Create dataset with one line per person with mileage score 
ageranef <- merge(ageranef, SCLdata, by = 'sub')
ageranef <- ageranef %>%
  dplyr::group_by(sub) %>%
  dplyr::slice(1) %>%
  dplyr::select(sub, age.c, intercept_age, slope_age)


# Person-specific implied total n-back effects for model accounting for age
ranef.age.pred.tot <- fixef(Model)[2] + # fixed effect for n-back
  fixef(Model)[12] * ageranef$age.c +   # fixed effect for n-back X age interaction # scale of age 
  ageranef$slope_age # estimate random effect of slopes


# relationship between the effect of N-back and age
age.nback.effect.sge_SCL <- ggplot(ageranef, aes(age.c, ranef.age.pred.tot)) + 
  geom_vline(xintercept = 0, size = .5, color = "black", linetype="solid") +
  geom_hline(yintercept = mean(ranef.age.pred.tot), size = .5, color = "black", linetype="solid") +
  geom_jitter(height = 0, width = 0, size = 4,
              shape = 21, colour = "black", fill = viridis(5)[3], alpha = .95, stroke = 1) + 
  stat_smooth(method = 'lm', ci = T, color = viridis(5)[3], size = 1) + 
  xlab("Age (mean centered)") +
  ylab(expression("Implied total heterogeneity of "~ SCL)) +
  #ylim(-.15, .1) +
  theme_bw() +
  theme(legend.position = "bottom", legend.direction = "horizontal", axis.title.x = element_text(size = 11), axis.text.x = element_text(size = 15), axis.title.y = element_text(size = 11), axis.text.y = element_text(size = 15), title = element_text(size = 18), legend.title = element_text(size = 12), legend.text = element_text(size = 12), legend.key = element_blank(), legend.key.width = unit(0.3, 'cm'), legend.key.size = unit(0.1, 'cm'), plot.title = element_text(size = 7.5, face = "bold"), strip.text = element_text(face = "bold", size = 10))
age.nback.effect.sge_SCL
```

### SDNN
```{r}
# scale the data
SDNNdata$age.c <- scale(SDNNdata$age, center = T, scale = F)

# fit model
library(lmerTest)
Model <- lmer(data = SDNNdata, SDNN~beforeafter*condition*age.c+(1+beforeafter|sub))
summary(Model)

### calculate the contribution of age 
# Get interaction coefficient between age and beforeafter
bXa.coeff <- fixef(Model)[12]  # the estimate parameter of age.c and beforeafter interaction effect

#Random effect in model for beforeafter - squared to compute variance 
residvalvar <- 0.01036

#Calculate the implied total random effect variance - interaction squared multiplied by variance in age + residual variance.
imptotalvalvar <- bXa.coeff^2*sd(SDNNdata$age.c)^2 + residvalvar


## Variance in mu difference explained by age
V_SDNN <- 1 - (residvalvar/imptotalvalvar)
V_SDNN

# plot the data
## computing random effects 
ageranef_original <- as.data.frame(ranef(Model)) %>% 
  dplyr::select("grp","term","condval")

library(reshape2)
ageranef <- dcast(ageranef_original, 
                  grp~term, 
                  timevar = c("condval")) 

ageranef <- ageranef %>% rename("intercept_age"="(Intercept)","slope_age"="beforeafteron","sub"="grp")


## Create dataset with one line per person with mileage score 
ageranef <- merge(ageranef, SDNNdata, by = 'sub')
ageranef <- ageranef %>%
  dplyr::group_by(sub) %>%
  dplyr::slice(1) %>%
  dplyr::select(sub, age.c, intercept_age, slope_age)


## Person-specific implied total n-back effects for model accounting for age
ranef.age.pred.tot <- fixef(Model)[2] + # fixed effect for n-back
  fixef(Model)[12] * ageranef$age.c +   # fixed effect for n-back X age interaction # scale of age 
  ageranef$slope_age # estimate random effect of slopes


## relationship between the effect of N-back and age
age.nback.effect.sge_SDNN <- ggplot(ageranef, aes(age.c, ranef.age.pred.tot)) + 
  geom_vline(xintercept = 0, size = .5, color = "black", linetype="solid") +
  geom_hline(yintercept = mean(ranef.age.pred.tot), size = .5, color = "black", linetype="solid") +
  geom_jitter(height = 0, width = 0, size = 4,
              shape = 21, colour = "black", fill = viridis(5)[3], alpha = .95, stroke = 1) + 
  stat_smooth(method = 'lm', ci = T, color = viridis(5)[3], size = 1) + 
  xlab("Age (mean centered)") +
  ylab(expression("Implied total heterogeneity of "~ SDNN)) +
  #ylim(-.15, .1) +
  theme_bw() +
  theme(legend.position = "bottom", legend.direction = "horizontal", axis.title.x = element_text(size = 11), axis.text.x = element_text(size = 15), axis.title.y = element_text(size = 11), axis.text.y = element_text(size = 15), title = element_text(size = 18), legend.title = element_text(size = 12), legend.text = element_text(size = 12), legend.key = element_blank(), legend.key.width = unit(0.3, 'cm'), legend.key.size = unit(0.1, 'cm'), plot.title = element_text(size = 7.5, face = "bold"), strip.text = element_text(face = "bold", size = 10))
age.nback.effect.sge_SDNN
```

### Breathing rate
```{r}
# scale the data
RESPdata$age.c <- scale(RESPdata$age, center = T, scale = F)

# fit model
Model <- lmer(data = RESPdata, RESP~beforeafter*condition*age.c+(1+beforeafter|sub))
summary(Model)

# calculate the contribution of age 
## Get interaction coefficient between age and beforeafter
bXa.coeff <- fixef(Model)[12] 

## Random effect in model for beforeafter - squared to compute variance 
residvalvar <- 0.002668

## Calculate the implied total random effect variance - interaction squared multiplied by variance in age + residual variance.
imptotalvalvar <- bXa.coeff^2*sd(RESPdata$age.c)^2 + residvalvar

## Variance in mu difference explained by age
V_RESP <- 1 - (residvalvar/imptotalvalvar)
V_RESP 

# plot the data
## computing random effects 
ageranef_original <- as.data.frame(ranef(Model)) %>% 
  dplyr::select("grp","term","condval")

library(reshape2)
ageranef <- dcast(ageranef_original, 
                  grp~term, 
                  timevar = c("condval")) 

ageranef <- ageranef %>% rename("intercept_age"="(Intercept)","slope_age"="beforeafteron","sub"="grp")


## Create dataset with one line per person with mileage score 
ageranef <- merge(ageranef, RESPdata, by = 'sub')
ageranef <- ageranef %>%
  dplyr::group_by(sub) %>%
  dplyr::slice(1) %>%
  dplyr::select(sub, age.c, intercept_age, slope_age)

## Person-specific implied total n-back effects for model accounting for age
ranef.age.pred.tot <- fixef(Model)[2] + # fixed effect for n-back
  fixef(Model)[12] * ageranef$age.c +   # fixed effect for n-back X age interaction # scale of age 
  ageranef$slope_age # estimate random effect of slopes

## relationship between the effect of N-back and age
age.nback.effect.sge_breathingrate <- ggplot(ageranef, aes(age.c, ranef.age.pred.tot)) + 
  geom_vline(xintercept = 0, size = .5, color = "black", linetype="solid") +
  geom_hline(yintercept = mean(ranef.age.pred.tot), size = .5, color = "black", linetype="solid") +
  geom_jitter(height = 0, width = 0, size = 4,
              shape = 21, colour = "black", fill = viridis(5)[3], alpha = .95, stroke = 1) + 
  stat_smooth(method = 'lm', ci = T, color = viridis(5)[3], size = 1) + 
  xlab("Age (mean centered)") +
  ylab(expression("Implied total heterogeneity of "~ breathing~rate)) +
  #ylim(-.15, .1) +
  theme_bw() +
  theme(legend.position = "bottom", legend.direction = "horizontal", axis.title.x = element_text(size = 11), axis.text.x = element_text(size = 15), axis.title.y = element_text(size = 11), axis.text.y = element_text(size = 15), title = element_text(size = 18), legend.title = element_text(size = 12), legend.text = element_text(size = 12), legend.key = element_blank(), legend.key.width = unit(0.3, 'cm'), legend.key.size = unit(0.1, 'cm'), plot.title = element_text(size = 7.5, face = "bold"), strip.text = element_text(face = "bold", size = 10))
age.nback.effect.sge_breathingrate
```

### PD
```{r}
# scale the data
PDdata$age.c <- scale(PDdata$age, center = T, scale = F)

# fit model
Model <- lmer(data = PDdata, PD~beforeafter*condition*age.c+(1+beforeafter|sub))
summary(Model)

# calculate the contribution of age 
# Get interaction coefficient between age and beforeafter
bXa.coeff <- fixef(Model)[12] 

# Random effect in model for beforeafter - squared to compute variance 
residvalvar <- 0.0003204

# Calculate the implied total random effect variance - interaction squared multiplied by variance in age + residual variance.
imptotalvalvar <- bXa.coeff^2*sd(PDdata$age.c)^2 + residvalvar

# Variance in mu difference explained by age
V_PD <- 1 - (residvalvar/imptotalvalvar)
V_PD

# plot the data
## computing random effects 
ageranef_original <- as.data.frame(ranef(Model)) %>% 
  dplyr::select("grp","term","condval")

library(reshape2)
ageranef <- dcast(ageranef_original, 
                  grp~term, 
                  timevar = c("condval")) 

ageranef <- ageranef %>% rename("intercept_age"="(Intercept)","slope_age"="beforeafteron","sub"="grp")


## Create dataset with one line per person with mileage score 
ageranef <- merge(ageranef, PDdata, by = 'sub')
ageranef <- ageranef %>%
  dplyr::group_by(sub) %>%
  dplyr::slice(1) %>%
  dplyr::select(sub, age.c, intercept_age, slope_age)



## Person-specific implied total n-back effects for model accounting for age
ranef.age.pred.tot <- fixef(Model)[2] + # fixed effect for n-back
  fixef(Model)[12] * ageranef$age.c +   # fixed effect for n-back X age interaction # scale of age 
  ageranef$slope_age # estimate random effect of slopes



## relationship between the effect of N-back and age
age.nback.effect.sge_pupildiameter <- ggplot(ageranef, aes(age.c, ranef.age.pred.tot)) + 
  geom_vline(xintercept = 0, size = .5, color = "black", linetype="solid") +
  geom_hline(yintercept = mean(ranef.age.pred.tot), size = .5, color = "black", linetype="solid") +
  geom_jitter(height = 0, width = 0, size = 4,
              shape = 21, colour = "black", fill = viridis(5)[3], alpha = .95, stroke = 1) + 
  stat_smooth(method = 'lm', ci = T, color = viridis(5)[3], size = 1) + 
  xlab("Age (mean centered)") +
  ylab(expression("Implied total heterogeneity of "~ pupil~diameter)) +
  #ylim(-.15, .1) +
  theme_bw() +
  theme(legend.position = "bottom", legend.direction = "horizontal", axis.title.x = element_text(size = 11), axis.text.x = element_text(size = 15), axis.title.y = element_text(size = 11), axis.text.y = element_text(size = 15), title = element_text(size = 18), legend.title = element_text(size = 12), legend.text = element_text(size = 12), legend.key = element_blank(), legend.key.width = unit(0.3, 'cm'), legend.key.size = unit(0.1, 'cm'), plot.title = element_text(size = 7.5, face = "bold"), strip.text = element_text(face = "bold", size = 10))
age.nback.effect.sge_pupildiameter
```
### PVRC
```{r}
# scale the data
PVRCdata$age.c <- scale(PVRCdata$age, center = T, scale = F)

# fit model
Model <- lmer(data = PVRCdata, PVRC~beforeafter*condition*age.c+(1+beforeafter|sub))
summary(Model)

# calculate the contribution of age 
# Get interaction coefficient between age and beforeafter
bXa.coeff <- fixef(Model)[12] 

# Random effect in model for beforeafter - squared to compute variance 
residvalvar <- 0.0002471

# Calculate the implied total random effect variance - interaction squared multiplied by variance in age + residual variance.
imptotalvalvar <- bXa.coeff^2*sd(PVRCdata$age.c)^2 + residvalvar

# Variance in mu difference explained by age
V_PVRC <- 1 - (residvalvar/imptotalvalvar)
V_PVRC 

# plot the data
## computing random effects 
ageranef_original <- as.data.frame(ranef(Model)) %>% 
  dplyr::select("grp","term","condval")

library(reshape2)
ageranef <- dcast(ageranef_original, 
                  grp~term, 
                  timevar = c("condval")) 

ageranef <- ageranef %>% rename("intercept_age"="(Intercept)","slope_age"="beforeafteron","sub"="grp")


## Create dataset with one line per person with mileage score 
ageranef <- merge(ageranef, PVRCdata, by = 'sub')
ageranef <- ageranef %>%
  dplyr::group_by(sub) %>%
  dplyr::slice(1) %>%
  dplyr::select(sub, age.c, intercept_age, slope_age)



## Person-specific implied total n-back effects for model accounting for age
ranef.age.pred.tot <- fixef(Model)[2] + # fixed effect for n-back
  fixef(Model)[12] * ageranef$age.c +   # fixed effect for n-back X age interaction # scale of age 
  ageranef$slope_age # estimate random effect of slopes



## relationship between the effect of N-back and age
age.nback.effect.sge_PVRC <- ggplot(ageranef, aes(age.c, ranef.age.pred.tot)) + 
  geom_vline(xintercept = 0, size = .5, color = "black", linetype="solid") +
  geom_hline(yintercept = mean(ranef.age.pred.tot), size = .5, color = "black", linetype="solid") +
  geom_jitter(height = 0, width = 0, size = 4,
              shape = 21, colour = "black", fill = viridis(5)[3], alpha = .95, stroke = 1) + 
  stat_smooth(method = 'lm', ci = T, color = viridis(5)[3], size = 1) + 
  xlab("Age (mean centered)") +
  ylab(expression("Implied total heterogeneity of "~ PVRC)) +
  #ylim(-.15, .1) +
  theme_bw() +
  theme(legend.position = "bottom", legend.direction = "horizontal", axis.title.x = element_text(size = 11), axis.text.x = element_text(size = 15), axis.title.y = element_text(size = 11), axis.text.y = element_text(size = 15), title = element_text(size = 18), legend.title = element_text(size = 12), legend.text = element_text(size = 12), legend.key = element_blank(), legend.key.width = unit(0.3, 'cm'), legend.key.size = unit(0.1, 'cm'), plot.title = element_text(size = 7.5, face = "bold"), strip.text = element_text(face = "bold", size = 10))
age.nback.effect.sge_PVRC
```

### Blink rate (BL)
```{r}
# scale the data
BLdata$age.c <- scale(BLdata$age, center = T, scale = F)

# fit model
Model <- lmer(data = BLdata, BL~beforeafter*condition*age.c+(1+beforeafter|sub))
summary(Model)

# calculate the contribution of age 
# Get interaction coefficient between age and beforeafter
bXa.coeff <- fixef(Model)[12] 

# Random effect in model for beforeafter - squared to compute variance 
residvalvar <- 2.304e-06

# Calculate the implied total random effect variance - interaction squared multiplied by variance in age + residual variance.
imptotalvalvar <- bXa.coeff^2*sd(BLdata$age.c)^2 + residvalvar

# Variance in mu difference explained by age
V_BL <- 1 - (residvalvar/imptotalvalvar)
V_BL

# plot the data
## computing random effects 
ageranef_original <- as.data.frame(ranef(Model)) %>% 
  dplyr::select("grp","term","condval")

library(reshape2)
ageranef <- dcast(ageranef_original, 
                  grp~term, 
                  timevar = c("condval"))

ageranef <- ageranef %>% rename("intercept_age"="(Intercept)","slope_age"="beforeafteron","sub"="grp")


## Create dataset with one line per person with mileage score 
ageranef <- merge(ageranef, BLdata, by = 'sub')
ageranef <- ageranef %>%
  dplyr::group_by(sub) %>%
  dplyr::slice(1) %>%
  dplyr::select(sub, age.c, intercept_age, slope_age)



## Person-specific implied total n-back effects for model accounting for age
ranef.age.pred.tot <- fixef(Model)[2] + # fixed effect for n-back
  fixef(Model)[12] * ageranef$age.c +   # fixed effect for n-back X age interaction # scale of age 
  ageranef$slope_age # estimate random effect of slopes



## relationship between the effect of N-back and age
age.nback.effect.sge_blinkrate <- ggplot(ageranef, aes(age.c, ranef.age.pred.tot)) + 
  geom_vline(xintercept = 0, size = .5, color = "black", linetype="solid") +
  geom_hline(yintercept = mean(ranef.age.pred.tot), size = .5, color = "black", linetype="solid") +
  geom_jitter(height = 0, width = 0, size = 4,
              shape = 21, colour = "black", fill = viridis(5)[3], alpha = .95, stroke = 1) + 
  stat_smooth(method = 'lm', ci = T, color = viridis(5)[3], size = 1) + 
  xlab("Age (mean centered)") +
  ylab(expression("Implied total heterogeneity of "~ blink~rate)) +
  #ylim(-.15, .1) +
  theme_bw() +
  theme(legend.position = "bottom", legend.direction = "horizontal", axis.title.x = element_text(size = 11), axis.text.x = element_text(size = 15), axis.title.y = element_text(size = 11), axis.text.y = element_text(size = 15), title = element_text(size = 18), legend.title = element_text(size = 12), legend.text = element_text(size = 12), legend.key = element_blank(), legend.key.width = unit(0.3, 'cm'), legend.key.size = unit(0.1, 'cm'), plot.title = element_text(size = 7.5, face = "bold"), strip.text = element_text(face = "bold", size = 10))
age.nback.effect.sge_blinkrate
```

## 3.6 Subjective ratings
### 3.6.1 Workload
```{r}
# read data
MWdata <- read_excel("/Users/betty/Desktop/Workload(processed).xlsx", sheet=1)

# fit model
Model <- lmer(data = MWdata, meanscore ~ agegroup2*condition*dimension + (1|sub))
summary(Model)
anova(Model)

# the post hoc analysis for the main effect of countermeasure type
pwc.workload.condition <- MWdata %>% pairwise_t_test(meanscore ~ condition, p.adjust.method = "bonferroni", paired = TRUE)
pwc.workload.condition 

# the post hoc analysis for the main effect of workload dimension
pwc.workload.dimension <- MWdata %>% pairwise_t_test(meanscore ~ dimension, p.adjust.method = "bonferroni", paired = TRUE) 
pwc.workload.dimension

# the simple simple effect analysis for the interaction effect of age group, countermeasure type, and workload dimension
## younger drivers (condition X dimension)
filter.age.Y <- filter(MWdata,agegroup2 =="Y") 
### fit model
Model <- lmer(data = filter.age.Y, meanscore~condition*dimension+(1|sub))
summary(Model)
anova(Model)
### pairwise comparisons of five countermeasure types under different workload dimensions
#### (1) mental
filter1.MWdata<- filter(filter.age.Y, dimension =="mental") 
library(lmerTest)
Model <- lmer(data = filter1.MWdata, meanscore~condition+(1|sub))
summary(Model)
anova(Model)
pwc <- filter1.MWdata %>% pairwise_t_test(meanscore ~ condition, p.adjust.method = "bonferroni", paired = TRUE) 
pwc
#### (2) physical
filter2.MWdata<- filter(filter.age.Y, dimension =="physical") 
Model <- lmer(data = filter2.MWdata, meanscore~condition+(1|sub))
summary(Model)
anova(Model)
pwc <- filter2.MWdata %>% pairwise_t_test(meanscore ~ condition, p.adjust.method = "bonferroni", paired = TRUE) 
pwc
#### (3) time
filter3.MWdata<- filter(filter.age.Y, dimension =="time") 
Model <- lmer(data = filter3.MWdata, meanscore~condition+(1|sub))
summary(Model)
anova(Model)
pwc <- filter3.MWdata %>% pairwise_t_test(meanscore ~ condition, p.adjust.method = "bonferroni", paired = TRUE) 
pwc
#### (4) effort
filter4.MWdata<- filter(filter.age.Y, dimension =="effort") 
Model <- lmer(data = filter4.MWdata, meanscore~condition+(1|sub))
summary(Model)
anova(Model)
pwc <- filter4.MWdata %>% pairwise_t_test(meanscore ~ condition, p.adjust.method = "bonferroni", paired = TRUE) 
pwc
#### (5) frustration
filter5.MWdata<- filter(filter.age.Y, dimension =="frustration") 
Model <- lmer(data = filter5.MWdata, meanscore~condition+(1|sub))
summary(Model)
anova(Model)
pwc <- filter5.MWdata %>% pairwise_t_test(meanscore ~ condition, p.adjust.method = "bonferroni", paired = TRUE) 
pwc
#### (6) performance
filter6.MWdata<- filter(filter.age.Y, dimension =="performance")
Model <- lmer(data = filter6.MWdata, meanscore~condition+(1|sub))
summary(Model)
anova(Model)
pwc <- filter6.MWdata %>% pairwise_t_test(meanscore ~ condition, p.adjust.method = "bonferroni", paired = TRUE) 
pwc



## middle-aged drivers (condition X dimension)
filter.age.M <- filter(MWdata,agegroup2 =="M") 
### fit model
Model <- lmer(data = filter.age.M, meanscore~condition*dimension+(1|sub))
summary(Model)
anova(Model)
### pairwise comparisons of five countermeasure types under different workload dimensions
#### (1) mental
filter1.MWdata<- filter(filter.age.M, dimension =="mental") 
Model <- lmer(data = filter1.MWdata, meanscore~condition+(1|sub))
summary(Model)
anova(Model)
pwc <- filter1.MWdata %>% pairwise_t_test(meanscore ~ condition, p.adjust.method = "bonferroni", paired = TRUE) 
pwc
#### (2) physical
filter2.MWdata<- filter(filter.age.M, dimension =="physical") 
Model <- lmer(data = filter2.MWdata, meanscore~condition+(1|sub))
summary(Model)
anova(Model)
pwc <- filter2.MWdata %>% pairwise_t_test(meanscore ~ condition, p.adjust.method = "bonferroni", paired = TRUE) 
pwc
#### (3) time
filter3.MWdata<- filter(filter.age.M, dimension =="time") 
Model <- lmer(data = filter3.MWdata, meanscore~condition+(1|sub))
summary(Model)
anova(Model)
pwc <- filter3.MWdata %>% pairwise_t_test(meanscore ~ condition, p.adjust.method = "bonferroni", paired = TRUE) 
pwc
#### (4) effort
filter4.MWdata<- filter(filter.age.M, dimension =="effort") 
Model <- lmer(data = filter4.MWdata, meanscore~condition+(1|sub))
summary(Model)
anova(Model)
pwc <- filter4.MWdata %>% pairwise_t_test(meanscore ~ condition, p.adjust.method = "bonferroni", paired = TRUE) 
pwc
#### (5) frustration
filter5.MWdata<- filter(filter.age.M, dimension =="frustration") 
Model <- lmer(data = filter5.MWdata, meanscore~condition+(1|sub))
summary(Model)
anova(Model)
pwc <- filter5.MWdata %>% pairwise_t_test(meanscore ~ condition, p.adjust.method = "bonferroni", paired = TRUE) 
pwc
#### (6) performance
filter6.MWdata<- filter(filter.age.M, dimension =="performance")
Model <- lmer(data = filter6.MWdata, meanscore~condition+(1|sub))
summary(Model)
anova(Model)
pwc <- filter6.MWdata %>% pairwise_t_test(meanscore ~ condition, p.adjust.method = "bonferroni", paired = TRUE) 
pwc
```


### 3.6.2 Acceptance
```{r}
# read data
USdata <- read_excel("/Users/betty/Desktop/Acceptance (processed).xlsx", sheet=1)

# as factors & groupby
USdata$condition <- as.factor(USdata$condition)
USdata$agegroup2 <- as.factor(USdata$agegroup2)
USdata %>%
  group_by(agegroup2,condition)

# Analysis of younger drivers
filter.Y<- filter(USdata,agegroup2 =="Y") 

## main effect of countermeasure type on usefulness
Us.fried.useful.young <- filter.Y %>% friedman_test(usefulness ~ condition |sub)
Us.fried.useful.young
## the post hoc analysis for the main effect of countermeasure type on usefulness
pwc.useful.young <- filter.Y %>%  wilcox_test(usefulness ~ condition, paired = TRUE, p.adjust.method = "bonferroni")
pwc.useful.young

## main effect of countermeasure type on satisfaction
Us.fried.satisfaction.young <- filter.Y %>% friedman_test(satisfaction ~ condition |sub)
Us.fried.satisfaction.young
## the post hoc analysis for the main effect of countermeasure type on satisfaction
pwc.satisfaction.young <- filter.Y %>%  wilcox_test(satisfaction ~ condition, paired = TRUE, p.adjust.method = "bonferroni")
pwc.satisfaction.young


# Analysis of middle-aged drivers
filter.M <- filter(USdata,agegroup2 =="M") 
## main effect of countermeasure type on usefulness
Us.fried.useful.middle <- filter.M %>% friedman_test(usefulness ~ condition |sub)
Us.fried.useful.middle
## the post hoc analysis for the main effect of countermeasure type on usefulness
pwc.useful.middle <- filter.M %>%  wilcox_test(usefulness ~ condition, paired = TRUE, p.adjust.method = "bonferroni")
pwc.useful.middle

## main effect of countermeasure type on satisfaction
Us.fried.satisfaction.middle <- filter.M %>% friedman_test(satisfaction ~ condition |sub)
Us.fried.satisfaction.middle
## the post hoc analysis for the main effect of countermeasure type on usefulness
pwc.satisfaction.middle <- filter.M %>%  wilcox_test(satisfaction ~ condition, paired = TRUE, p.adjust.method = "bonferroni")
pwc.satisfaction.middle
```
